---
title: 'Peer-graded Assignment: Course Project 1'
output:
  html_document:
    df_print: paged
---

```{r}
library(knitr)
opts_chunk$set(echo= FALSE)
```


# Loading and preprocessing the data
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
                    destfile="activity_file")
unzip("activity_file")
activity <- read.csv("activity.csv")
str(activity)
```


# What is mean total number of steps taken per day?

## Calculate the total number of steps taken per day

```{r}
daily_steps <- setNames(aggregate(activity$steps, list(activity$date), sum, na.rm=T),
                        c("Date", "Steps"))
daily_steps
```


## Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
ggplot(daily_steps, aes(Steps))+
    geom_histogram(bins = 15)+
    labs(x="Total daily steps", title = "Histogram of total daily steps")
```

## Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(daily_steps$Steps)
```

The mean total number of step taken per day is 9,354 and the median is 10,935.

# What is the average daily activity pattern?
```{r}
avg_steps <- setNames(aggregate(activity$steps, 
                                list(activity$interval), mean, na.rm=T),
                      c("Interval", "Avg_Steps"))

ggplot(avg_steps, aes(Interval, Avg_Steps))+
    geom_line()+
    labs(x= "Time interval (min)",
         y="Average steps", title = "Average steps taken over 5 min intervals")
```

#Which 5-minute interval contains the maximum number of steps?
```{r}
avg_steps$Interval[which.max(avg_steps$Avg_Steps)]
```

The 835 min interval contains the maximum number of steps

# Imputing missing values

## Calculate and report the total number of missing values in the dataset
```{r}
missing <- subset(activity, is.na(steps))
nrow(missing)
```


## Devise a strategy for filling in all of the missing values in the dataset
```{r}
new_activity <- merge(avg_steps, activity, by.x = "Interval", by.y = "interval")
for (i in 1:nrow(new_activity)){
    if (is.na(new_activity$steps[[i]])){
        new_activity$steps[[i]] <- ceiling(new_activity$Avg_Steps[[i]])
    }
}
```


## Create a new dataset  with the missing data filled in.
```{r}
library(dplyr)
new_activity <- new_activity %>%
                    select(steps, date, Interval) %>%
                    arrange(date)
```


## Make a histogram of the total number of steps taken each day
```{r}
ggplot(new_activity, aes(steps))+
    geom_histogram(bins = 15)+
    labs(x="Total daily steps", title = "Histogram of new total daily steps")
```


## Calculate and report the mean and median total number of steps taken per day
```{r}
new_daily_steps <- setNames(aggregate(new_activity$steps, list(new_activity$date), 
                                      sum, na.rm=T), c("Date", "Steps"))
new_daily_steps

summary(new_daily_steps$Steps)
```

The new mean daily steps is 10,785 and the new median is 10,909. Both the mean and the median increased in value and got closer to each other.

# Are there differences in activity patterns between weekdays and weekends?

## Create a new factor variable indicating whether a given date is a weekday or weekend day.

```{r}
new_activity$date <- as.Date(new_activity$date)
new_activity$weekday <- weekdays(new_activity$date)
new_activity$weekday <- factor(new_activity$weekday,
                               levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"),
                               labels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

for (i in 1:nrow(new_activity)){
    if (new_activity$weekday[[i]] == "Saturday" | new_activity$weekday[[i]] == "Sunday") {
    new_activity$factordays[[i]] <- 1
    } else {
        new_activity$factordays[[i]] <- 0
    }
}

new_activity$factordays <- factor(new_activity$factordays, levels = c(0, 1), labels = c("Weekday", "Weekend"))
```


## time series plot  of the 5-minute interval (x-axis) and the average number of steps taken by day type
```{r}
weekend <- subset(new_activity, factordays== "Weekend")
week <- subset(new_activity, factordays == "Weekday")


avg_steps$weekend <- setNames(aggregate(weekend$steps, 
                                list(weekend$Interval), mean, na.rm=T),
                      c("Interval", "Avg_Steps"))
avg_steps$week <- setNames(aggregate(week$steps, 
                                         list(week$Interval), mean, na.rm=T),
                               c("Interval", "Avg_Steps"))


weekend <- ggplot(avg_steps, aes(Interval, weekend$Avg_Steps))+
    geom_line(col="blue")+
    labs(x= "Time interval (min)",
         y="Average steps", title = "Average steps taken over 5 min intervals (weekend)")

week <- ggplot(avg_steps, aes(Interval, week$Avg_Steps))+
    geom_line(col="red")+
    labs(x= "Time interval (min)",
         y="Average steps", title = "Average steps taken over 5 min intervals (weekend)")

library(ggpubr)

ggarrange(weekend, week, ncol=1)

```


